---
title: "Russian_Analysis"
author: "Elizaveta Volkova"
date: "11/04/2020"
output:
  html_document:
    df_print: paged
---
```{r,message=F}
library(dplyr)
library(ggplot2)
library(forecast)
library(tseries)
library(dplyr, quietly = T)
require(timelineR)
require(astsa)
library(tidyr)
```
```{r}
cases<-read.csv("covid19-russia-cases.csv")
tests<-read.csv("tests_updated.csv")
```

```{r}
#head(cases)
#head(tests)
#Updated date column to recognizable by R format in tests file
#tests$Date <- strptime(as.character(tests$Date), "%d.%m.%Y")
#tests$Date<- format(tests$Date, "%Y-%m-%d")

#Counted change of tests compared to previous day
tests<-tests%>%mutate(changeFromPrevious = Tests - lag(Tests))
#The first day of tests information provided is 2020-02-21 thus I substract the rows with NAs
tests<-tests[24:nrow(tests),]
#test<-select(tests,Date,changeFromPrevious,Tests)
tests
```

## Made dataframe with number of cases groupped by each date

I substract all the dates before 2020-03-05 as the previous information seems to be unreliable - the total number of confirmed cases fluctuates between [1:3] (that might be affected by Diamond Princess cases that were not actually in Russia) and becomes 1 again on 2020-03-05.
```{r}
cases_by_date<-cases%>%group_by(Date)%>%summarise(sum(Confirmed),sum(Deaths),sum(Recovered),sum(Day.Confirmed))
#nrow(cases_by_date)
#cases_by_date<-dplyr::bind_rows(cases_by_date, test)
#tail(cases_by_date)
ts_confirmed=ts(cases_by_date$`sum(Confirmed)`)
#ts.plot(ts_confirmed)
#axis(1, cases_by_date$Date, format(cases_by_date$Date), cex.axis = .7)
#nrow(cases_by_date)
cases_by_date$Date<-as.POSIXct(cases_by_date$Date)
cases_by_date$`sum(Day.Confirmed)`<-as.numeric(cases_by_date$`sum(Day.Confirmed)`)
cases_by_date$`sum(Confirmed)`<-as.numeric(cases_by_date$`sum(Confirmed)`)
cases_by_date$`sum(Deaths)`<-as.numeric(cases_by_date$`sum(Deaths)`)
tests$Tests<-as.numeric(tests$Tests)
tests$Date<-as.POSIXct(tests$Date)
tests$changeFromPrevious<-as.numeric(tests$changeFromPrevious)
```
```{r}
cases_by_date<-cases_by_date[5:nrow(cases_by_date),]
```
## Timeseries plot for Cumulative number of confirmed cases
```{r}
data_cols=c("sum(Confirmed)")
titles=c("sum(Confirmed)"="Cumulative number of confirmed cases")
ylabels=c("sum(Confirmed)"="Confirmed cases")

plot_grob <- plot_timeline(cases_by_date,data_cols = data_cols,titles=titles,ylabels=ylabels)
```



## Timeseries plot for Cumulative number of deaths
```{r}
data_cols=c("sum(Deaths)")
titles=c("sum(Deaths)"="Cumulative number of deaths")
ylabels=c("sum(Deaths)"="Deaths")
plot_grob <- plot_timeline(cases_by_date,data_cols = data_cols,titles=titles,ylabels=ylabels)
```

## Timeseries plot for Cumulative number of tests
```{r}
data_cols=c("Tests")
titles=c("Tests"="Cumulative number of tests")
ylabels=c("Tests"="Tests")
plot_grob <- plot_timeline(tests,data_cols = data_cols,titles=titles,ylabels=ylabels)
```


## Histogram of daily confirmed cases
```{r}
ggplot(cases,aes(Date,Day.Confirmed))+geom_bar(stat="identity")+theme(axis.text.x = element_text(size = 10,hjust = 0,angle = -70)) +ggtitle("Daily confirmed cases")+xlab("Date")+ylab("Confirmed cases")
```

## Histogram of daily tests
```{r}
ggplot(tests,aes(Date,changeFromPrevious))+geom_bar(stat="identity")+theme(axis.text.x = element_text(size = 10,hjust = 0,angle = -70))+ggtitle("Daily number of tests")+xlab("Date")+ylab("Tests completed")
```


## Auto ARIMA for cases groupped by date
```{r}

prediction_time<-30#Prediction period (days)

auto.arima(cases_by_date$`sum(Confirmed)`)
auto<-arima(cases_by_date$`sum(Confirmed)`, order=c(1,2,0))

prediction<-forecast(auto,h=prediction_time)
dates<-c(as.Date(cases_by_date$Date),as.Date(seq(as.Date(cases_by_date$Date[nrow(cases_by_date)]),by="day",length.out = prediction_time+1)))
#dates<-as.POSIXct(dates)
prediction

plot(prediction,xaxt='n',main=("Forecast of total number of cases"))
at  <-  seq(1,nrow(cases_by_date)+prediction_time,length.out=12)
axis(1, at = at+5, labels = dates[at],cex.axis = .7,las=3)  
```

## Auto ARIMA for deaths groupped by date
```{r}

prediction_time<-30#Prediction period (days)

auto.arima(cases_by_date$`sum(Deaths)`)
auto<-arima(cases_by_date$`sum(Deaths)`, order=c(1,2,3))

prediction<-forecast(auto,h=prediction_time)
dates<-c(as.Date(cases_by_date$Date),as.Date(seq(as.Date(cases_by_date$Date[nrow(cases_by_date)]),by="day",length.out = prediction_time+1)))
#dates<-as.POSIXct(dates)

plot(prediction,xaxt='n',main=("Forecast of total number of deaths"))
at  <-  seq(1,nrow(cases_by_date)+prediction_time,length.out=12)
axis(1, at = at+5, labels = dates[at],cex.axis = .7,las=3)  
```

## Auto ARIMA for cases by date
```{r}

prediction_time<-30#Prediction period (days)

auto.arima(cases_by_date$`sum(Day.Confirmed)`)
auto<-arima(cases_by_date$`sum(Day.Confirmed)`, order=c(0,2,3))

prediction<-forecast(auto,h=prediction_time)
dates<-c(as.Date(cases_by_date$Date),as.Date(seq(as.Date(cases_by_date$Date[nrow(cases_by_date)]),by="day",length.out = prediction_time+1)))
#dates<-as.POSIXct(dates)
prediction

plot(prediction,xaxt='n',main=("Forecast for daily confirmed cases"))
at  <-  seq(1,nrow(cases_by_date)+prediction_time,length.out=12)
axis(1, at = at+5, labels = dates[at],cex.axis = .7,las=3)  
```

## Normalization of number of cases by number of tests
```{r}
cases_by_date$Date<-as.factor(cases_by_date$Date)
for_merge<-select(tests,Date,Tests)
for_merge$Date<-as.factor(for_merge$Date)
normalized<-merge(x = cases_by_date, y = for_merge, by = "Date", all.x = TRUE)
normalized$normalization<-ifelse(is.na(normalized$Tests),normalized$`sum(Confirmed)`,normalized$`sum(Confirmed)`/normalized$Tests)
#normalized<-select(normalized,Date,normalization)
normalized$Date<-as.POSIXct(normalized$Date)
normalized<-normalized[1:nrow(normalized)-1,]
```
## Timeseries plot for Cumulative number of cases normalized by cumulative number of tests
```{r,eval=T}
data_cols=c("normalization")
titles=c("normalization"="Cumulative number of cases normalized by cumulative number of tests")
ylabels=c("normalization"="normalization")
plot_grob <- plot_timeline(normalized,data_cols,titles=titles,ylabels=ylabels)
```
## Auto ARIMA for cases by date
```{r}

prediction_time<-30#Prediction period (days)

auto.arima(normalized$normalization)
auto<-arima(normalized$normalization, order=c(0,2,1))

prediction<-forecast(auto,h=prediction_time)
dates<-c(as.Date(cases_by_date$Date),as.Date(seq(as.Date(cases_by_date$Date[nrow(cases_by_date)]),by="day",length.out = prediction_time+1)))
#dates<-as.POSIXct(dates)

plot(prediction,xaxt='n',main=("Forecast for Cumulative number of cases normalized by cumulative number of tests"))
at  <-  seq(1,nrow(cases_by_date)+prediction_time,length.out=12)
axis(1, at = at+5, labels = dates[at],cex.axis = .7,las=3)  
```
## Normalization of number of cases by number of tests
```{r}
cases_by_date$Date<-as.factor(cases_by_date$Date)
for_merge<-select(tests,Date,Tests,changeFromPrevious)
for_merge$Date<-as.factor(for_merge$Date)
normalized<-merge(x = cases_by_date, y = for_merge, by = "Date", all.x = TRUE)
normalized$normalization<-ifelse(is.na(normalized$Tests),normalized$`sum(Confirmed)`,normalized$`sum(Confirmed)`/normalized$Tests)
normalized<-select(normalized,Date,normalization)
normalized$Date<-as.POSIXct(normalized$Date)
normalized<-normalized[1:nrow(normalized)-1,]
normalized
```
## Timeseries plot for Cumulative number of cases normalized by cumulative number of tests
```{r,eval=T}
data_cols=c("normalization")
titles=c("normalization"="Cumulative number of cases normalized by cumulative number of tests")
ylabels=c("normalization"="normalization")
plot_grob <- plot_timeline(normalized,data_cols=data_cols,titles=titles,ylabels=ylabels)
```

## Auto ARIMA for cases by date
```{r}

prediction_time<-30#Prediction period (days)

auto.arima(normalized$normalization)
auto<-arima(normalized$normalization, order=c(0,2,1))

prediction<-forecast(auto,h=prediction_time)
dates<-c(as.Date(cases_by_date$Date),as.Date(seq(as.Date(cases_by_date$Date[nrow(cases_by_date)]),by="day",length.out = prediction_time+1)))
#dates<-as.POSIXct(dates)

plot(prediction,xaxt='n',main=("Forecast for Cumulative number of cases normalized by cumulative number of tests"))
at  <-  seq(1,nrow(cases_by_date)+prediction_time,length.out=12)
axis(1, at = at+5, labels = dates[at],cex.axis = .7,las=3)  
```

## Normalization of Number of daily cases by daily number of tests
```{r}
#cases_by_date
cases_by_date$Date<-as.factor(cases_by_date$Date)
for_merge<-select(tests,Date,Tests,changeFromPrevious)
for_merge$Date<-as.factor(for_merge$Date)
#for_merge
normalized<-merge(x = cases_by_date, y = for_merge, by = "Date", all.x = TRUE)
#normalized$normalization<-normalized$`sum(Day.Confirmed)`/normalized$changeFromPrevious
normalized$normalization<-ifelse(normalized$changeFromPrevious==0,NA,normalized$`sum(Day.Confirmed)`/normalized$changeFromPrevious)
#normalized$normalization<-ifelse(is.na(normalized$changeFromPrevious),NA,normalized$`sum(Day.Confirmed)`/normalized$changeFromPrevious)
normalized<-select(normalized,Date,normalization)
normalized$Date<-as.POSIXct(normalized$Date)
normalized<-normalized[1:nrow(normalized)-1,]
normalized<-normalized %>% drop_na()
#normalized
```
## Timeseries plot for number of daily cases by number of daily number of tests
```{r,eval=T}
data_cols=c("normalization")
titles=c("normalization"="Number of daily cases by daily number of tests")
ylabels=c("normalization"="normalization")
plot_grob <- plot_timeline(normalized,data_cols,titles=titles,ylabels=ylabels)
```

## Auto ARIMA for cases by date
```{r}

prediction_time<-30#Prediction period (days)

auto.arima(normalized$normalization)
auto<-arima(normalized$normalization, order=c(0,1,1))

prediction<-forecast(auto,h=prediction_time)
dates<-c(as.Date(cases_by_date$Date),as.Date(seq(as.Date(cases_by_date$Date[nrow(cases_by_date)]),by="day",length.out = prediction_time+1)))
#dates<-as.POSIXct(dates)

plot(prediction,xaxt='n',main=("Forecast for Number of daily cases by daily number of tests"))
at  <-  seq(1,nrow(cases_by_date)+prediction_time,length.out=12)
axis(1, at = at+5, labels = dates[at],cex.axis = .7,las=3)  
```

# Not needed part:

```{r}
seque<-NULL
seque$Date<-(seq(as.Date(cases_by_date$Date[nrow(cases_by_date)]),by="day",length.out = prediction_time+1))
seque<-as.data.frame(seque)
seque<-seque$Date[2:nrow(seque)]
for_out<-NULL
for_out$Date<-as.factor(seque)
for_out$point<-as.numeric(prediction$mean)
for_out$Lo80<-prediction$lower[,1]
for_out$Lo90<-prediction$lower[,2]
for_out$Hi80<-prediction$upper[,1]
for_out$Hi90<-prediction$upper[,2]

for_out<-as.data.frame(for_out)
#for_out

for_join<-select(for_out,Date,Lo80,Lo90,Hi80,Hi90)
#for_join

full<-NULL
full$Date<-(cases_by_date$Date)
full$point<-cases_by_date$`sum(Confirmed)`
full<-as.data.frame(full)
full<-rbind(full,for_out[c(1,2)])
full<-merge(x = full, y = for_join, by = "Date", all.x = TRUE)
full$Date<-as.POSIXct(full$Date)

plot(prediction,xaxt='n')
at  <-  seq(1,nrow(cases_by_date)+prediction_time,length.out=10)
axis(1, at = at+5, labels = full$Date[at],cex.axis = .7,las=3)  
```

```{r}
c(as.Date(cases_by_date$Date),as.Date(seq(as.Date(cases_by_date$Date[nrow(cases_by_date)]),by="day",length.out = prediction_time+1)))

typeof(full$Date)
```



```{r,eval=F}
seque<-NULL
seque$Date<-(seq(as.Date(cases_by_date$Date[nrow(cases_by_date)]),by="day",length.out = 21))
seque<-as.data.frame(seque)
seque<-seque$Date[2:nrow(seque)]
for_out<-NULL
for_out$Date<-as.factor(seque)
for_out$point<-as.numeric(prediction$mean)
for_out$Lo80<-prediction$lower[,1]
for_out$Lo90<-prediction$lower[,2]
for_out$Hi80<-prediction$upper[,1]
for_out$Hi90<-prediction$upper[,2]

for_out<-as.data.frame(for_out)
for_out

for_join<-select(for_out,Date,Lo80,Lo90,Hi80,Hi90)
for_join
```

```{r,eval=FALSE}
full<-NULL
full$Date<-(cases_by_date$Date)
full$point<-cases_by_date$`sum(Confirmed)`
full<-as.data.frame(full)
full<-rbind(full,for_out[c(1,2)])
full<-merge(x = full, y = for_join, by = "Date", all.x = TRUE)
full$Date<-as.POSIXct(full$Date)
full
data_cols=c("point")
#overlapping_plot_names <- list("point"=c("point","Lo80","Hi80"))
plot_grob <- plot_timeline(full,data_cols=data_cols)
```

```{r,eval=FALSE}
data_cols=c("changeFromPrevious")
plot_grob <- plot_timeline(tests,data_cols = data_cols)
```

https://cran.rstudio.com/web/packages/timelineR/vignettes/plot_timeline.html



```{r,eval=F}
data_cols=c("sum(Day.Confirmed)")
titles=c("sum(Day.Confirmed)"="Daily number of confirmed cases")
ylabels=c("sum(Day.Confirmed)"="Confirmed cases")

plot_grob <- plot_timeline(cases_by_date,data_cols = data_cols,titles=titles,ylabels=ylabels)
```
```{r,eval=FALSE}
data_cols=c("sum(Day.Confirmed)")
titles=c("sum(Day.Confirmed)"="Number of confirmed cases")
ylabels=c("sum(Day.Confirmed)"="Confirmed")
ylimits=list("sum(Day.Confirmed)"=c(0,150))
plot_grob <- plot_timeline(cases_by_date,data_cols = data_cols,titles=titles,ylabels=ylabels,ylimits = ylimits)
```

```{r,eval=FALSE}
prediction$fitted
prediction$upper
inds <- seq(as.Date(cases_by_date$Date[1]), as.Date(cases_by_date$Date[nrow(cases_by_date)]), by = "day")
plot(prediction, xaxt = "n")    # no x-axis 
Axis(inds, side = 1,
     at = seq(inds[1], tail(inds, 1) + 60, by = "3 months"),
     format = "%b %Y",cex.axis = .7)
#data_cols=c("Point Forecast")
#plot_grob <- plot_timeline(prediction,data_cols = data_cols)
#plot(cases[c(1,4)])
#1, cases_by_date$Date, format(cases_by_date$Date), cex.axis = .7
#ts.plot(cases$Day.Confirmed,prediction$pred, log = "y", lty = c(1,3))
```



```{r,eval=FALSE}
library(forecast)
require(tseries); require(astsa)
#plot(cases$Confirmed~cases$Date)
auto.arima(cases$Day.Confirmed)
auto<-arima(cases$Day.Confirmed, order=c(5,1,0))
prediction<-predict(auto,n.ahead=60)
prediction<-forecast(auto,h=60)
head(prediction)
plot(prediction)
data_cols=c("Point Forecast")
#plot_grob <- plot_timeline(prediction,data_cols = data_cols)
#plot(cases[c(1,4)])

#ts.plot(cases$Day.Confirmed,prediction$pred, log = "y", lty = c(1,3))
```
```{r,eval=FALSE}
auto.arima(cases$Confirmed)
auto<-arima(cases$Confirmed, order=c(5,1,0))
prediction<-predict(auto,n.ahead=60)
#prediction<-forecast(auto,h=60)
head(prediction)
plot(prediction)
data_cols=c("Point Forecast")
#plot_grob <- plot_timeline(prediction,data_cols = data_cols)
#plot(cases[c(1,4)])

#ts.plot(cases$Day.Confirmed,prediction$pred, log = "y", lty = c(1,3))
```
https://www.analyticsvidhya.com/blog/2016/02/time-series-forecasting-codes-python/ 



