---
title: "Project_583_Canada"
author: "Sofia Bahmutsky"
date: "16/04/2020"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### 1.

Loading the Canada data, goes up to April 24
```{r}
library(ggplot2)
library(dplyr)
library(timelineR)
library(data.table)
library(forecast)
library(tidyr)
library(gdata)
library(stringr)
```

```{r}
### second data set
cases_2 <- read.csv("/Users/Sofia/Desktop/Data583Project/canada-covid19/cases_2.csv", stringsAsFactors = F)
cases_2 <- subset(cases_2, Country.Region=='Canada')
#cases_2 <- cases_2[,-c(3:82)]
cases_2 <- cases_2[,-c(2, 3, 4)]

data_long <- gather(cases_2, date_report, n, X1.22.20:X4.24.20, factor_key=TRUE)
data_long$date_report <- as.character(data_long$date_report)

data_long$date_report<-sub("X", "", data_long$date_report)
data_long$date_report<-str_replace_all(data_long$date_report, "[.]", "/")
data_long$date_report <- as.Date(data_long$date_report, "%m/%d/%y")
data_long$date_report <- as.Date(data_long$date_report, "%d/%m/%Y")
```

## ARIMA for daily cases 
## Time series of Daily Cases - Canada wide
```{r}
#daily cases

canada <- data_long %>% group_by(data_long$date_report)
canada <- data_long %>%group_by(data_long$date_report) %>% summarise(total_cases= sum(n))
#canada <- subset(canada[-1,])
new <- diff(canada$total_cases, differences = 1)
new <- c(0, new)
canada$new <- new
setnames(canada, old = c('data_long$date_report'), new = c('date'))

ts_cases <- ts(canada$new)
ts.plot(ts_cases, xlab="Days", ylab="Daily New Cases")

acf(diff(canada$new))
plot(diff(canada$new))

arima_cases <- auto.arima(canada$new, seasonal = F)
arima_cases

arima_cases2 <- arima(canada$new, order=c(2, 1, 3))
arima_cases2

prediction_time <- 30
preds <- forecast(arima_cases2, h=30)

dates<-c(as.Date(canada$date),as.Date(seq(as.Date(canada$date[nrow(canada)]),by="day",length.out = prediction_time+1)))
a <-plot(preds,xaxt='n',main=("Forecast of daily confirmed cases - Canada"), ylab="Number of New Cases")
at  <-  seq(1,nrow(canada)+prediction_time,length.out=12)
axis(1, at = at+5, labels = dates[at],cex.axis = .7,las=3) 
```


## Daily cases for Canada - Normalized
```{r}
tests <- read.csv("/Users/Sofia/Desktop/Data583Project/canada-covid19/testing_cumulative.csv", stringsAsFactors = F)

tests <- select(tests, date_testing, cumulative_testing)
setnames(tests, old = c('date_testing', 'cumulative_testing'), new = c('Date', 'csum'))
tests$Date <- as.Date(tests$Date, "%d-%m-%Y")

tests <- tests %>%group_by(Date) %>% summarise(cumulative_tests= sum(csum))

new_tests <- diff(tests$cumulative_tests, differences = 1)

new_tests <- c(0, new_tests)
tests$daily_tests <- new_tests
setnames(tests, old = c('Date'), new = c('date'))

canada
canada_normal <- inner_join(canada, tests, by = 'date')

canada_normal$proportion <- canada_normal$new/canada_normal$daily_tests
##############################################################

ts_cases_normalized <- ts(canada_normal$proportion)
ts.plot(ts_cases_normalized, xlab="Days", ylab="Positive daily cases (normalized)")

canada_normal <- drop_na(canada_normal)
acf(diff(canada_normal$proportion, 2))

arima_cases_normal <- auto.arima(canada_normal$proportion, seasonal = F)
arima_cases_normal

arima_cases_normal2 <- arima(canada_normal$proportion, order=c(7, 2, 2))
arima_cases_normal2


prediction_time <- 30
preds2 <- forecast(arima_cases_normal2, h=30)

dates<-c(as.Date(canada_normal$date),as.Date(seq(as.Date(canada_normal$date[nrow(canada_normal)]),by="day",length.out = prediction_time+1)))
plot(preds2,xaxt='n',main=("Forecast of Normalized Cases (proportion)"), ylab="Positive daily cases (normalized)")
at  <-  seq(1,nrow(canada_normal)+prediction_time,length.out=12)
axis(1, at = at+3, labels = dates[at],cex.axis = .7,las=3)

```














#### Not using this part ...
## Canada-wide Cumulative
```{r}
#cumulative cases
cumulative <- aggregate(data_long$n, by=list(Category=data_long$date_report), FUN=sum)

ts_cases <- ts(cumulative$x)
ts.plot(ts_cases, xlab="Days", ylab="Cumulative Cases")

```

## ARIMA for cumulative (I'm not sure if this is useful...I thought ARIMA was to observe patterns in the daily data, cumulative data shows entirely different pattern. )

```{r}
#daily cases
arima_cases_cumulative <- auto.arima(cumulative$x, seasonal = F)
arima_cases_cumulative

preds_cumulative <- forecast(arima_cases_cumulative, h=30)
plot(preds_cumulative)

```



## Province Breakdown - Cumulative
```{r}
# cases by province
cases_prov <- aggregate(data_long$n, by=list(Category=data_long$date_report, data_long$Province.State), FUN=sum)
DT <- data.table(cases_prov, key = "Group.2")
DT <- DT[, csum := cumsum(x), by = key(DT)]
setnames(DT, old = c('Category','Group.2'), new = c('date','province'))

ggplot(data=DT,
       aes(x=DT$date, y=DT$csum, colour=DT$province)) +
       geom_line()


DT_reduced <- subset(DT, DT$province == c('Alberta', 'British Columbia', 'Ontario', 'Quebec'))

ggplot(data=DT_reduced,
       aes(x=DT_reduced$date, y=DT_reduced$csum, colour=DT_reduced$province)) +
       geom_line()
```